Investigating DNA d

### 2020-09-26 manuscript.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{TBA}

@@

\textcolor{purple}{Could add more here about all-atom simulations?}

@@

\begin{align}
 	\mathscr{C}_{00}=&\Expect{\chi(x_t)\chi(x_t)^\intercal}_t\\
 	\mathscr{C}_{01}=&\Expect{\chi(x_t)\chi(x_{t+\tau})^\intercal}_t\\
 	\mathscr{C}_{11}=&\Expect{\chi(x_{t+\tau})\chi(x_{t+\tau})^\intercal}_{t+\tau}\\
	\\
 	R_2[\chi]=&\norm{\mathscr{C}_{00}^{-1/2}\mathscr{C}_{01}\mathscr{C}_{11}^{-1/2}}_F^2 +1
 	%R_2^{val}[\chi]=&\norm{\mathscr{(C_{00}^{val})}^{-1/2}\mathscr{C}_{01}^{val}\mathscr({C}_{11}^{val})^{-1/2}}_F^2 + 1
\end{align}\label{CK1}

@@

We compared scores for the permutation-free distances, intermolecular, and intra+intermolecular distances and found small differences between the three that varied by sequences (show all sequences in supplemental), but overall we did not observe a loss in kinetic information when using the reduced feature set. We found a substantial increase in VAMP-2 score when using reciprocal pairwise distances and chose reciprocal permutation-free coordinates as a consistent feature set. These features were normalized and passed into sequence-specific SRVs.

@@

In our analysis, we found that the AT-all sequence, given its repetitive structure and lack of GC-content, produced the cleanest dynamics and displayed a canonical spectral gap between modes. For this reason, we lead our discussion with this sequence and use it as a case study to work through our SRV-MSM pipeline step-by-step. Our first task was to identify the SRV lag time that was longer than the intrinsic Markov timescales of the system, yet short enough to resolve the dynamics of interest \citep{Phys2018MarkovValidation}. We found that most implied timescales converge at an SRV lag time of 600 ps. We kept a looser constraint on the convergence of the leading mode -- corresponding to the overall hybridization/dissociation process -- as it tends to have a longer transit period and becomes more difficult to fulfill Markovian conditions. It should be noted that our lag time selection was informed by the other sequences as well to maintain consistency. Next we selected an optimal number of SRV components to include in our analysis. After a certain point, higher order dynamical modes provide diminishing contributions the overall kinetic variance as measured by the Vamp-2 score, and the model can begin fitting on statistical noise in the trajectory data instead of the true dynamics \citep{McGibbon2015VariationalKinetics}. It is also more difficult to perform kmeans clustering on a high dimensional space, especially when those higher dimensions are less kinetically relevant \citep{Pande2010EverythingAsk}. For these reasons, the number of slow SRV components should be carefully selected based on the specific system of interest. As shown in (Figure \ref{fig:srv_crossval-AT-all}), we see diminishing returns in the VAMP-2 score after five slow modes and select these modes as our optimized SRV basis. Next, we seek to interpret the physical relevance of these leading modes by plotting the Pearson correlation of each mode with the 100 intermolecular distances between strands. The quantitative meaning of these coordinates can be difficult to interpret given their nonlinear relationship to the SRV collective variables, but the relative different between these correlations shows which coordinates are most affected by (or effective on) each process. For example, the first slow mode shows a significant positive correlation to each distance and the strongest correlation with matching base pair distances (shown along the main diagonal). Given these relationships and the substantially longer timescale of this process, we can deduce that this leading mode corresponds to the dynamics of the overall hybridization and dissociation process. The next four SRV components all show a relatively high correlation along different offset diagonals. These diagonals correspond to the intermolecular distances between shifted base pairs and point to the existence of gradually faster shifting or "slithering" processes. Slow slithering mechanisms have similarly been reported in simulation studies to occur on orders of magnitude longer timescales than underlying fast dynamics such as fraying \citep{Markegard2015, Xiao2019, Maciejczyk2014DNAModel}.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=50mm, 
        scale=0.5]{Figs/figs_0804/srv_crossval-AT-all.png}
        \caption{5-fold cross validation procedure to select number of SRV coordinates. We look for both the converge of the VAMP-2 and consistent scores between folds to ensure that the model is not fitting on statistical noise. For AT-all we observed a spectral gap after the 5th coordinate. Cross-validation on other sequences is in the SI. (Add AT-all feature scores here as well?)}
        \label{fig:srv_crossval-AT-all}
	\end{center}
\end{figure}

@@

From this analysis we can determine that the slowest dynamics are fully characterized by the hybridization process and slithering behavior of out of register base pairs. Although this is qualitatively informative, we can access a more holistic picture of sequence kinetics and thermodynamics by using these SRV coordinates as a basis on which to construct an MSM. Because these coordinates are already capturing a majority of the system's kinetic variance, they serve as an ideal basis on which to group frames into microstates. We performed k-means clustering, and optimized the number of microstates at 200 by monitoring VAMP-2 score. Next, we selected an MSM lag time in a similar fashion to our SRV lag time selection process. In Figure \ref{fig:AT-all_dynamic} we compare the convergence of SRV-MSM and TICA-MSM timescales, where SRV-MSM timescales converge consistently faster and to higher values of all leading modes. This enables us to select a shorter lag time and build a higher resolution model than we could from an analogous TICA basis. Using an MSM lag time of 1.2 ns, we then built a Bayesian MSM to calculate transition probability matrix between each microstate. Finally, PCCA+ spectral clustering was implemented to group these microstates into macrostates that each represent a collection of metastable structures. Previous work have used a common set of microstates and/or performed manual clustering of microstates based on physical read outs from simulation data (stacking score, energies, etc) \citep{Pinamonti2017PredictingModels, PinamontiTheModels}. Although these techniques are useful for performing comparisons between sequences, we saw better results when optimizing MSMs to capture the most detail of sequence individually and thus developed an independent set of microstates and macrostates for each sequence. For AT-all, we kept to the convention of clustering into n+1 macrostates, where n is the number of slow components captured by the MSMs. To visualize these six macrostates, we project the data into the two leading TICA coordinates. Although SRVs outperform these coordinates for the purpose of MSM construction, TICA represent good high variance collective variables on which to visualize free energies and state assignments \citep{Sidky}. It is clear in Figure \ref{fig:AT-all_in_one} that the PCCA state assignments are capturing free energy minima in TICA space as independent states. After assigning these macrostates we can then calculate their relative probabilities and free energies, visualize representative molecular renderings, and estimates transition probabilities between states. In this final step, we use a minimum flux cutoff of 2e-6 in order to mitigate erroneous quick transition or skipping between states.

@@

In our coarse-grained SRV-MSM, we observe an "aligned hybridized" state, dissociated state, and four "shifted" states characterized by different combinations of out-of-register base pairings. These shifted states consist of 2 or 4 base pair shifting (single or doubled shifted) in either the 5' or 3' direction with varied stability and transition fluxes between kinetically neighboring states. This state decomposition is expected given the repetitive nature of the AT-all sequence and previous computational results that find shifted conformations form "deep kinetic traps" \citep{Xiao2019, Phys2014}. We found a substantial difference in thermodynamic stability and kinetic behavior between the 5'A shifted states and 3'T shifted states. This difference can be accounted for by examining experimental studies on the thermodynamics of "dangling ends" -- unpaired bases adjacent to the paired duplex -- and "inert tails" -- free bases that extend beyond the dangling end \citep{Michele2014EHybridization}. Dangling 5' ends have a consistently stabilizing effect, and inert tails tend to decrease stability as they increase in length, particularly at lower ionic strengths. It has been shown that 5' dangling ends with one inert tail have higher melting temperatures and are energetically favorable compared to 3' ends \citep{Senior1988InfluenceDuplexes, Dickman2012ThermodynamicDNAs}. This is likely due to 5' ends preferentially stacking on the core duplex compared to 3' ends and/or 3' ends binding few counterions and perturbing the duplex structure and solvation environment \citep{Doktycz1990ThermodynamicATGC}. Furthermore, DNA nearest neighbor (NN) calculations that include dangling end contributions (Figure \ref{fig:AT-all_in_one}) predict that the specific nearest neighbors bonds in the 5' shifted state are more energetically favorable than those in the 3' shifted state. Although NN calculations were accurate in predicting the intact thermodynamics behavior of these sequences, they assume all-or-nothing hybridization and do not take inert tails into account \citep{Santalucia2004TM, SantaLucia1998AThermodynamics} Taken together, these experimental results substantiate why we observe the single-shifted 5' state to be only slightly less stable relative to the aligned hybridized state. Furthermore, we expect implied timescales of the two leading modes to be on the same order of magnitude as these reflect the average "dwell" time between transition into and out of these states. The stability of the single-shifted 3' state is notably lower than that of the single-shifted 5' state. In fact, the free energy is about equal to that of the double-shifted 5' state, despite it having two less intact base pairs. This is partially explained by nearest neighbor calculations, however these do not account for the differential effects of the 5' and 3' inert tails or kinetic considerations which we will discuss more below. The double-shifted 3' state is more stable than one might expect given the difference in stability between the single and double 5' shifted states. This might be attributed to the asymptotic effects of inert tail destabilization -- longer tails tend to have a diminishing effect on overall strand stability \citep{Michele2014EHybridization}. Accordingly, the longer inert tails in the double-shifted states have less of an effect on relative 5' vs. 3' stability compared to the shorter tails in the single-shifted states.

@@

For all proceeding sequences, we use the same SRV and MSM lag times, number of microstates, and cross-validation procedures as in the AT-all case. Although these might not represent fully optimized hyperparameters for each sequence, they preserve Markovian properties of the system and provide adequate resolution to evaluate macrostates \textcolor{magenta}{not sure if this undermines the point I make earlier about building unique microstates and macrostates for each sequence so as to optimize the model}. The number of SRV coordinates and MSM macrostates are varied based on the kinetics and thermodynamics behavior of specific sequences.

@@

For our analysis, we consider C:T and G:A base pairs in the GC-ends shifted states as non-interacting dangling ends such that each shifted conformation has four total dangling ends and two inert tails. Although internal base pair mismatches can cause substantial conformational distortions such as kinking, terminal mismatches have been shown to be slightly stabilizing and have a minimal effect on helical character \citep{Santalucia2004TM, DiMichele2014EffectHybridization}. In the context of the 3spn2 model, these ends are accounted for via intra-strand base stacking and inter-strand cross-stacking interactions \citep{Hinckley2013AnHybridization}. The only direct interaction between non-WC basepairs is parameterized by isotropic excluded volume potential, which is likely much more simplistic than the true mismatch interaction. Despite fewer intact base pairs, we find these structures stabilizing enough to account for the relatively small population of conformations we observed in these shifted states (Figure \ref{fig:GC-end_in_one}). Contrary to AT-all thermodynamics, nearest neighbor calculations predict the 3' shifted states to be slightly more stable than 5' shifted states, again without taking into account entropic or enthalipic (electrostatic) effects from inert tails \textcolor{Purple}{Message here should be that inert tails, looping, other effects have more substantial contribution than two-state NN calculation alone}. However, state populations reveal that the 5' single-shifted is again significantly more stable than the 3'. Interestingly, we found that 5' single-shifted stability might be elevated by a substantial portion of conformations retaining one intact GC base pair. Visualizations reveal that the shifted oligos -- particularly in the 5' shifted state -- have a tendency to sacrifice some helical conformational entropy in a way that facilitates G:C termini bonding even when all available A:T bonds are formed out-of-register.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Figs/figs_0804/GC-end_in_one.PNG}
        \caption{a) Thermo value calculated from Santalucia data and from MSMs b) TICA projections of freen energies and state clustering c) and transitions probabilities between states \textcolor{magenta}{Need to fix 5' double shifted sequence, make same adjustments as for AT-all above}}
        \label{fig:GC-end_in_one}
	\end{center}
\end{figure}

@@

To compare these single-shifted 5' prime states with the corresponding AT-all state, we employ diffusion maps built on an equal sampling of 5000 conformations from 5' shifted state of both sequences. Diffusion maps generate a low dimensional embedding based on some metric for diffusive distance and are well-suited to find subtle structural differences in temporally disconnected data \citep{Coifman2006DiffusionMaps, Ferguson2010SystematicMaps}. We used all 100 intermolecular distances (as opposed to the 55 permutation free coordinates used to construct SRV-MSM) as our distance metric, making it easier to discern structures that form on either identical end of the shifted conformation. This created a degenerate 2nd and 3rd diffusion modes, with nearly equal eigenvalues, differentiating between looping at the identical "top" and "bottom" of the strands. In Figure \ref{fig:GC-end_dmaps} we present the first two non-trivial diffusion map eigenfunctions and show representations of the degenerate third coordinate in the SI (Figure \ref{fig:GC-end_dmaps_full}. Diffusion maps built from samples of the 3' shifted states are also shown in the SI. The first diffusion mode clearly delineates between the GC-end and AT-all shifted conformations and correlates highly with the average distance between the 3' end and its shifted complementary pair. This reveals that the mismatch C:T-pairs are never bound -- a consequence of the excluded volume interaction \textcolor{purple}{state explicitly that C and T have greater excluded volume than G and A in 3spn.2}-- whereas the AT-all pairs are mostly bound with occasional fraying indicated by small AT-all overlap in the GC-end region. This effect may be exaggerated given that C and T base pairs are assigning slightly higher exluded volume radii in 3spn2. The second diffusion mode, which correlates highly with the average distance between 3' and 5' ends, has higher values for GC-end than AT-all. Because the GC-end termini do not bind out of register, we find that they are readily able to form stabilizing contacts despite the shifted conformation of the duplex as a whole. These "shifted-loop" bonds are shown to be uniquely stable for GC-end conformations in the 5' shifted state, and their existence in the simulations confirmed by molecular renderings of these regions Although AT-all shifted ends tend to stay bound out-of-register, the second diffusion coordinate shows some population of inert tails that fold back onto the helix.

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Figs/figs_0804/GC-core_tracking_modes.png}
        \caption{Show how SRV coordinates correspond to physical features over time and during sample dissociation and hybridization events. Should probably go to SI if we also include GC-core diffusion maps \textcolor{magenta}{These single trajectory figures have proven useful for visualizing SRV dynamics, could be employed for GC-mix comparisons as well}}
        \label{fig:GC-core_tracking_modes}
	\end{center}
\end{figure}

@@

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Figs/figs_0804/GC-mix_transitions.PNG}
        \caption{WC base pair distance and molecular renderings along two representative GC-mix hybridization and dissociation events.}
        \label{fig:GC-mix_transitions}
	\end{center}
\end{figure}

@@

@@

For GC-core, each base pair and cutoff combination appears linear with temperature, indicating a barrierless and diffusion-driven process. As expected, the acceleration factor is also highly dependent on the choice of parameters, but, given that terminal fraying likely dominates the signal, we approximate that fraying dynamics are accelerated 100-300x relative to experiments. In contrast, GC-end is distinctly exponential, especially at base pairs closer to the core of the duplex. This behavior makes sense given the thermodynamic barrier to G:C fraying, but it also deviates the GC-end experimental linear fit. In addition to fraying, we notice that the rate of slithering increases at higher temperature. Moreover, the thermodynamic stability of these shifted conformations increases relative to that of intact duplex as temperature increases. For higher T-jumps, it is possible that a significant portion of the starting ensemble is in a shifted state (cite Ryan's work) and that the fast dynamics are increasingly influenced by slower "looping" G:C bonding that we observed in our diffusion map analysis above. It is less likely that shifting mechanisms would have their own distinct signals, given that these occur at far lower frequency than fraying or full-dissociation, but we believe that further experimental work is necessary to investigate this phenomenon.

The temperature-dependent analysis can also inform out interpretation of the 3spn2 model. Previous work has shows that kinetic association rates were accelerated by about 200x relative to experiment \citep{Hinckley2013AnHybridization}. This seems to be on the same order of magnitude as frayed acceleration, but dissociation is noticeably less accelerated when compared to experiment. It is not surprising that different dynamics might be accelerated at different rates, especially considering that 3spn2 was not parameterized based on any dynamical properties \citep{?}.

@@

@@

\begin{enumerate}
	\item Does not explicitly pick up on nucleation/zippering events. Because these are all unique sites (and zippering dynamics could vary at each site) it would be difficult to properly sample and interpret the unique corresponding mode. Furthermore these zippering dynamics occur significantly faster than slithering or GC-core fraying and cannot be captured via our save rate. 
	\item Coarse grain-models makes it difficult to compare timescales. Different degrees of freedom might have different accelerations.
	\item Limited by memory (number of frames) required to capture equilibrium trajectories at a save rate that enabled adequate resolution for the dynamics of interest.
	\item A smoother free energy surface minimizes the time spent in these intermediate states, and our featurization method is not sensitive to certain kinds of dynamics
	\item Sensitive to 3spn2 artefacts. Slithering may be overexpressed (and is in fact suppressed in \citep{Romano2013DNADependence}, but similar results have been shown in all-atom studies as well \citep{Xiao2019}. Base-pair mismatch interactions may be somewhat unphysical and alter state populations. Although we present slithering as a possible explanation for GC-end signal, there has been no conclusive evidence of these dynamics outside out all-atom and coarse-grained simulation studies. We hope that these data will present a

\end{enumerate}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]{Investigating DNA d}

@@

\textcolor{purple}{Could add more background on here about all-atom simulations, but these rarely explore fully hybridization/dissocition events}

@@

\begin{align*}
 	\mathscr{C}_{00}=&\Expect{\chi(x_t)\chi(x_t)^\intercal}_t\\
 	\mathscr{C}_{01}=&\Expect{\chi(x_t)\chi(x_{t+\tau})^\intercal}_t\\
 	\mathscr{C}_{11}=&\Expect{\chi(x_{t+\tau})\chi(x_{t+\tau})^\intercal}_{t+\tau}\\
	\\
 	R_2[\chi]=&\norm{\mathscr{C}_{00}^{-1/2}\mathscr{C}_{01}\mathscr{C}_{11}^{-1/2}}_F^2 +1
 	%R_2^{val}[\chi]=&\norm{\mathscr{(C_{00}^{val})}^{-1/2}\mathscr{C}_{01}^{val}\mathscr({C}_{11}^{val})^{-1/2}}_F^2 + 1
\end{align*}\label{CK1}

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=80mm,
        scale=0.5]{Figs/figs_0804/all_permute_vamps_grouped.png}
        \caption{5-fold cross validation to calculate the VAMP-2 score for each feature set. The inverse distances showed improvement across sequences, and the 55 reduced coordinate set performed about as well as larger features set.}
        \label{fig:allseq_features_vamp2}
	\end{center}
\end{figure}

We compared scores for the permutation-free distances, intermolecular, and intra+intermolecular distances and found small differences between the three that varied by sequences (show all sequences in supplemental), but overall we did not observe a loss in kinetic information when using the reduced feature set (\ref{fig:allseq_features_vamp2}). We found a substantial increase in VAMP-2 score when using reciprocal pairwise distances and chose reciprocal permutation-free coordinates to train the model. The smaller feature set enabled faster training times and better statistics over permutatable distances, without appearing to suffer a loss in generality or model resolution. These features were normalized and passed into sequence-specific SRVs.

@@

In our analysis, we found that the AT-all sequence, given its repetitive structure and lack of GC-content, produced the cleanest dynamics and displayed a canonical spectral gap between modes. For this reason, we lead our discussion with this sequence and use it as a case study to work through our SRV-MSM pipeline step-by-step. Our first task was to identify the SRV lag time that was longer than the intrinsic Markov timescales of the system, yet short enough to resolve the dynamics of interest \citep{Phys2018MarkovValidation}. We found that most implied timescales converge at an SRV lag time of 1.2 ns. We kept a looser constraint on the convergence of the leading mode -- corresponding to the overall hybridization/dissociation process -- as it tends to have a longer transit period and becomes more difficult to fulfill Markovian conditions. It should be noted that our lag time selection was informed by the other sequences as well to maintain consistency. Next we selected an optimal number of SRV components to include in our analysis. After a certain point, higher order dynamical modes provide diminishing contributions the overall kinetic variance as measured by the Vamp-2 score, and the model can begin fitting on statistical noise in the trajectory data instead of the true dynamics \citep{McGibbon2015VariationalKinetics}. It is also more difficult to perform kmeans clustering on a high dimensional space, especially when those higher dimensions are less kinetically relevant \citep{Pande2010EverythingAsk}. For these reasons, the number of slow SRV components should be carefully selected based on the specific system of interest. As shown in (Figure \ref{fig:allseq_srv_crossval}), we see diminishing returns in the VAMP-2 score after five slow modes and select these modes as our optimized SRV basis. Next, we seek to interpret the physical relevance of these leading modes by plotting the Pearson correlation of each mode with the 100 intermolecular distances between strands. The quantitative meaning of these coordinates can be difficult to interpret given their nonlinear relationship to the SRV collective variables, but the relative different between these correlations shows which coordinates are most affected by (or effective on) each process. For example, the first slow mode shows a significant positive correlation to each distance and the strongest correlation with matching base pair distances (shown along the main diagonal). Given these relationships and the substantially longer timescale of this process, we can deduce that this leading mode corresponds to the dynamics of the overall hybridization and dissociation process. The next four SRV components all show a relatively high correlation along different offset diagonals. These diagonals correspond to the intermolecular distances between shifted base pairs and point to the existence of gradually faster "slithering" processes. Slow slithering mechanisms have similarly been reported in simulation studies to occur on orders of magnitude longer timescales than underlying fast dynamics such as fraying \citep{Markegard2015, Xiao2019, Maciejczyk2014DNAModel}.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=120mm, 
        scale=0.5]{Figs/figs_0804/allseq_cross_val_scores_lag-40.png}
        \caption{5-fold cross validation procedure to select number of SRV coordinates. We look for the converge of the VAMP-2 scores, inconsistent scores between folds, and deviation between training and test data as indicators that the model has begun fitting on statistical noise.}
        \label{fig:allseq_srv_crossval}
	\end{center}
\end{figure}

@@

From this analysis we can determine that the slowest dynamics are fully characterized by the association/dissociation process and slithering behavior of out of register base pairs. Although this is qualitatively informative, we can access a more holistic picture of sequence kinetics and thermodynamics by using these SRV coordinates as a basis on which to construct an MSM. Because these coordinates are already capturing a majority of the system's kinetic variance, they serve as an ideal basis on which to group frames into microstates. We performed k-means clustering, and optimized the number of microstates at 200 by monitoring VAMP-2 score. Next, we selected an MSM lag time in a similar fashion to our SRV lag time selection process. In Figure \ref{fig:AT-all_dynamic} we compare the convergence of SRV-MSM and TICA-MSM timescales, where SRV-MSM timescales converge consistently faster and to higher values of all leading modes. This enables us to select a shorter lag time and build a higher resolution model than we could from an analogous TICA basis. Setting the MSM lag time to the same 1.2 ns we used for our SRVs, we built a Bayesian MSM to calculate transition probability matrix between each microstate. Finally, PCCA+ spectral clustering was implemented to group these microstates into macrostates that each represented a collection of metastable structures. Previous works have used a common set of microstates and/or performed manual clustering of microstates based on physical read outs from simulation data (stacking score, energies, etc) \citep{Pinamonti2017PredictingModels, PinamontiTheModels}. Although these techniques are useful for performing comparisons between sequences, we saw better results when optimizing MSMs to capture the most detail of sequence individually and thus developed an independent set of microstates and macrostates for each sequence. For AT-all, we kept to the convention of clustering into n+1 macrostates, where n is the number of slow components captured by the MSMs. To visualize these six macrostates, we project the data into the two leading TICA coordinates. Although SRVs outperform these coordinates for the purpose of MSM construction, TICA coordinates represent good high variance collective variables on which to visualize free energies and state assignments \citep{Sidky}. It is clear in Figure \ref{fig:AT-all_in_one} that the PCCA state assignments are capturing free energy minima in TICA space as independent states. After assigning these macrostates we can then calculate their relative probabilities and free energies, visualize representative molecular renderings, and estimates transition probabilities between states. In this final step, we use a minimum flux cutoff of 2e-6 in order to mitigate erroneous quick transition or skipping between states.

@@

In our coarse-grained SRV-MSM, we observe an "aligned hybridized" state, dissociated state, and four "shifted" states characterized by different combinations of out-of-register base pairings. These shifted states consist of 2 or 4 base pair shifting (single or doubled shifted) in either the 5' or 3' direction with varied stability and transition fluxes between kinetically neighboring states. This state decomposition is expected given the repetitive nature of the AT-all sequence and previous computational results that find shifted conformations form "deep kinetic traps" \citep{Xiao2019, Phys2014}. We found a substantial difference in thermodynamic stability and kinetic behavior between the 5'A shifted states and 3'T shifted states. This difference can be accounted for by examining experimental studies on the thermodynamics of "dangling ends" -- unpaired bases adjacent to the paired duplex -- and "inert tails" -- free bases that extend beyond the dangling end \citep{Michele2014EHybridization}. Dangling 5' ends have a consistently stabilizing effect, and inert tails tend to decrease stability as they increase in length, particularly at lower ionic strengths. It has been shown that 5' dangling ends with one inert tail have higher melting temperatures and are energetically favorable compared to 3' ends \citep{Senior1988InfluenceDuplexes, Dickman2012ThermodynamicDNAs}. This is might be attributed to some combination of 5' tails preferentially stacking on the core duplex and 3' tails perturb the duplex structure and solvation environment \citep{Doktycz1990ThermodynamicATGC}. Furthermore, DNA nearest neighbor (NN) calculations that include dangling end contributions (Figure \ref{fig:AT-all_in_one}) predict that the specific nearest neighbors bonds in the 5' shifted state are more energetically favorable than those in the 3' shifted state. Although NN calculations were accurate in predicting the intact thermodynamics behavior of these sequences, they assume all-or-nothing hybridization and do not take inert tails into account \citep{Santalucia2004TM, SantaLucia1998AThermodynamics} Taken together, these experimental results substantiate why we observe the single-shifted 5' state to be only slightly less stable relative to the aligned hybridized state. Furthermore, we expect implied timescales of the two leading modes to be on the same order of magnitude as these reflect the average "dwell" time between transition into and out of these states. The stability of the single-shifted 3' state is notably lower than that of the single-shifted 5' state. In fact, the free energy is about equal to that of the double-shifted 5' state, despite it having two less intact base pairs. This is partially explained by nearest neighbor calculations, however these do not account for the differential effects of the 5' and 3' inert tails or kinetic considerations which we will discuss more below. The double-shifted 3' state is more stable than one might expect given the difference in stability between the single and double 5' shifted states. This might be attributed to the asymptotic effects of inert tail destabilization -- longer tails tend to have a diminishing effect on overall strand stability \citep{Michele2014EHybridization}. Accordingly, the longer inert tails in the double-shifted states have less of an effect on relative 5' vs. 3' stability compared to the shorter tails in the single-shifted states.

@@

For all proceeding sequences, we use the same SRV and MSM lag times, number of microstates, and cross-validation procedures as in the AT-all case. For each sequence, these preserve Markovian properties of the system and provide some consistency to compare across models. The number of SRV coordinates and MSM macrostates are determined based on cross-validation and the kinetic behavior of specific sequences.

@@

For our analysis, we consider C:T and G:A base pairs in the GC-ends shifted states as non-interacting dangling ends such that each shifted conformation has four total dangling ends and two inert tails. Although internal base pair mismatches can cause substantial conformational distortions such as kinking, terminal mismatches have been shown to be slightly stabilizing and have a minimal effect on helical character \citep{Santalucia2004TM, DiMichele2014EffectHybridization}. In the context of the 3spn2 model, these ends are accounted for via intra-strand base stacking and inter-strand cross-stacking interactions \citep{Hinckley2013AnHybridization}. The only direct interaction between non-WC basepairs is parameterized by isotropic excluded volume potential, which is likely much more 3 primestic than the true mismatch interaction. Despite fewer intact base pairs, we find these structures stabilizing enough to account for the relatively small population of conformations we observed in these shifted states (Figure \ref{fig:GC-end_in_one}). Contrary to AT-all thermodynamics, nearest neighbor calculations predict the 3' shifted states to be slightly more stable than 5' shifted states, again without taking into account entropic or enthalpic effects from inert tails. However, state populations reveal that the 5' single-shifted state is again significantly more stable than the 3', indicating that dangling end and inert tail effects outweigh nearest neighbor effects alone. Interestingly, we found that 5' single-shifted stability might be elevated by a substantial portion of conformations retaining one intact GC base pair. Visualizations reveal that the shifted oligos -- particularly in the 5' shifted state -- have a tendency to sacrifice some helical conformational entropy in a way that facilitates G:C termini bonding even when all available A:T bonds are formed out-of-register.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Figs/figs_0804/GC-end_in_one.PNG}
        \caption{a) Thermo value calculated from Santalucia data and from MSMs b) TICA projections of freen energies and state clustering c) and transitions}
        \label{fig:GC-end_in_one}
	\end{center}
\end{figure}

@@

To compare these single-shifted 5' prime states with the corresponding AT-all state, we employ diffusion maps built on an equal sampling of 5000 conformations from the 5' single-shifted state of both sequences. Diffusion maps generate a low dimensional embedding based on some metric for diffusive distance and are well-suited to find subtle structural differences in temporally disconnected data \citep{Coifman2006DiffusionMaps, Ferguson2010SystematicMaps}. We used all 100 intermolecular distances (as opposed to the 55 permutation free coordinates used to construct SRV-MSM) as our distance metric, making it easier to discern structures that form on either identical end of the shifted conformation. This created a degenerate 2nd and 3rd diffusion modes, with nearly equal eigenvalues, differentiating between looping at the identical "top" and "bottom" of the strands. In Figure \ref{fig:GC-end_dmaps} we present the first two non-trivial diffusion map eigenfunctions and show representations of the degenerate third coordinate in the SI (Figure \ref{fig:GC-end_dmaps_full}). Diffusion maps built from samples of the 3' shifted states are also shown in the SI (Figure \ref{fig:GC-end_dmaps_3prime}). The first diffusion mode clearly delineates between the GC-end and AT-all shifted conformations and correlates highly with the average distance between the 3' end and its shifted complementary pair. This reveals that the mismatch C:T-pairs are never bound -- a consequence of the excluded volume interaction -- whereas the AT-all pairs are mostly bound with occasional fraying indicated by small AT-all overlap in the GC-end region. This effect may be exaggerated given that C and T base pairs are assigned slightly higher excluded volume radii in 3spn2 \citep{Hinckley2013AnHybridization}. The second diffusion mode, which correlates highly with the average distance between 3' and 5' ends, has higher values for GC-end than AT-all. Because the GC-end termini do not bind out of register, we find that they are readily able to form stabilizing contacts despite the shifted conformation of the duplex as a whole. These "shifted-loop" bonds are shown to be uniquely stable for GC-end conformations in the 5' shifted state, and their existence in the simulations confirmed by molecular renderings of these regions Although AT-all shifted ends tend to stay bound out-of-register, the second diffusion coordinate shows some population of inert tails that fold back onto the helix.

@@

\begin{figure} %[ht!]
	\begin{center}
        \includegraphics[width=120mm]{Figs/figs_0804/GC-core_tracking_modes.png}
        \caption{Show how SRV coordinates correspond to physical features over time and during sample dissociation and hybridization events. Should probably go to SI if we also include GC-core diffusion maps.}
        \label{fig:GC-core_tracking_modes}
	\end{center}
\end{figure}

@@

\begin{figure} %[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Figs/figs_0804/GC-mix_transitions.PNG}
        \caption{WC base pair distance and molecular renderings along two representative GC-mix hybridization and dissociation events.}
        \label{fig:GC-mix_transitions}
	\end{center}
\end{figure}

@@

@@

\begin{figure}  %[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Figs/figs_0804/fast-slow_end-core.png}
        \caption{Fitting temperature-dependent trends to the "slow" dissociation mode and "fast" fraying mode detected in experiments. Both show an exponential relationship between dissociation and temperature, however fraying appears more linear for GC-core than for GC-end.}
        \label{fig:relaxation-comparison}
	\end{center}
\end{figure}

@@

For GC-core, each base pair and cutoff combination appears linear with temperature, indicating a barrierless and diffusion-driven process. As expected, the acceleration factor is also highly dependent on the choice of parameters, but, given that terminal fraying likely dominates the signal, we approximate that fraying dynamics are accelerated by about two order of magnitude relative to experiments. In contrast, GC-end is distinctly exponential, especially at base pairs closer to the core of the duplex. This behavior makes sense given the thermodynamic barrier to G:C fraying, but it also deviates the GC-end experimental linear fit. In addition to fraying, we notice that the rate of slithering increases at higher temperature. Moreover, the thermodynamic stability of these shifted conformations increases relative to that of intact duplex. For higher T-jumps, it is possible that some portion of the starting ensemble is in a shifted state (cite Ryan's work) and that the fast dynamics are increasingly influenced by alternative "looping" G:C bonding that we observed in our diffusion map analysis above. It is less likely that shifting mechanisms would have their own distinct signals, given that these occur at far lower frequency than fraying or full-dissociation, but we believe that further experimental work is necessary to investigate this phenomenon.

The temperature-dependent analysis can also inform out interpretation of the 3spn2 model. Previous work has shows that kinetic association rates were accelerated by about 200x relative to experiment \citep{Hinckley2013AnHybridization}. This seems to be on the same order of magnitude as frayed acceleration, but dissociation is noticeably less accelerated when compared to experiment. It is not surprising that different dynamics might be accelerated at different rates, especially considering that 3spn2 was not parameterized based on any dynamical properties \citep{?}. Furthermore, more data is needed for explicit ion model comparisons.

@@

\section{\label{sec:conc}Conclusion}
\subsection{\label{sec:Results}Will look a lot like the last paragraph of the intro/abstract}

@@

\begin{enumerate}
	\item Does not explicitly pick up on nucleation/zippering events. Because these are all unique sites (and zippering dynamics could vary at each site) it would be difficult to properly sample and interpret the unique corresponding modes. Furthermore these zippering dynamics occur significantly faster than slithering or GC-core fraying and cannot be captured via our save rate. 
	\item Coarse grain-models makes it difficult to compare timescales. Different degrees of freedom might have different accelerations.
	\item Limited by memory (number of frames) required to capture equilibrium trajectories and rarity of hybridization/dissociation events
	\item A smoother free energy surface minimizes the time spent in these intermediate states
	\item Sensitive to 3spn2 artefacts. Slithering may be overexpressed (and is in fact suppressed in \citep{Romano2013DNADependence}, but similar results have been shown in all-atom studies as well \citep{Xiao2019}. Base-pair mismatch interactions may be somewhat unphysical and alter state populations. Although we present slithering as a possible explanation for GC-end signal, there has been no conclusive evidence of these dynamics outside out all-atom and coarse-grained simulation studies.
\end{enumerate}
